Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers

### 2024-03-30 2_SSL_NE/article/article_main.tex
various updates

Improvements in technology cost and performance, which in this study we generally call “innovations”, can occur via different mechanisms, including (but not limited to) learning by doing \cite{WRIGHT_1936, Arrow_1962}, targeted research and development (R\&D) efforts, or economies of scale \cite{johansson2012global, iea2020perspectives}. The role of these mechanisms at different stages of the technology life cycle, from research and development to demonstration, market formation and diffusion \cite{grubler2012policies}, is an area of active research \cite{kavlak2018evaluating,Ziegler2021}. Different mechanisms can result in different kinds of innovations, characterized by what changes in the technology (e.g., its architecture, materials, components, or manufacturing process) and the scale of resulting improvements (e.g., radical or incremental) \cite{Acemoglu_2022}. For example, R\&D efforts (sometimes also called “learning-by-researching”) \cite{Castelnuovo_2005} more often than other mechanisms can result in substantial or even radical improvements in technology performance. On the other hand, learning-by-doing is typically associated with continuous incremental improvements in the technology resulting from accumulated experience in technology demonstration, manufacturing, and use. Both mechanisms, as well as economies of scale, can also lead to radical (i.e., discontinuous) or incremental reductions in technology costs.

@@

We find that total LED device efficiency of such LEDs increased from 5.8\% to 38.8\% between 2003 and 2020. A majority of this increase can be attributed to a series of innovations in white LED technology resulting from R\&D efforts, with at least 8.5\% of the cumulative increase attributed specifically to technological innovations driven by technology spillovers. Our interviews also revealed a lesser but still important role of incremental manufacturing process improvements most likely resulting from learning-by-doing. However, substantial further research is needed to quantify the relative contribution of learning-by-doing vis-a-vis technological innovations to this progress. On the other hand, we find that innovations driven by technology spillovers contributed to nearly all consumer experience improvements in white LED lighting technology, thus playing an important role in widespread adoption of SSL in general illumination applications. Finally, our bottom-up LED manufacturing cost model shows a 95.5\% decrease in the cost of producing low-to-mid-power classic-chip-architecture GaN-based phosphor-converted warm-white LEDs from 1.11\$ to 0.05\$ (in 2020 USD) between 2003 and 2020. In contrast with LED performance improvements, where progress has been driven mostly by R\&D efforts, the dramatic decline in LED manufacturing cost has been a product of learning-by-doing resulting in higher yields across manufacturing steps and economies of scale resulting from increases in the sapphire wafer size. We provide the details of our findings for the progress in white LED efficiency, cost, and consumer experience, as well as the role of technology spillovers in this progress, in the remaining sections below.

@@

In \cref{fgr:breakthroughs_efficiency} we also compare, for the first time, efficiency improvements across sub-efficiencies over time, contrasting them with the physical limits of the corresponding loss channels. Since we focus on a best case scenario, and because we lack reliable information about the distribution of input parameters, which would require us to arbitrarily assign range values for inputs, these estimates do not show uncertainty ranges. Thus, our point estimates should be interpreted as the expected values in a best case scenario. We find that there has been consistent progress across all device sub-efficiencies in the recorded period. Specifically, between 2003 and 2020, forward voltage efficiency increased from 70\% to 99.5\%, internal quantum efficiency from 55\% to 90\%, electrical droop from 65\% to 90\%, light extraction efficiency from 60\% to 90\%, spectral efficiency from 74\% to 83\%, conversion efficiency (red) from 11\% to 45\%, conversion efficiency (green) from 19\% to 61\%. Notably, some sub-efficiencies for the most recent devices considered in our study are now within $\sim10\%$ of their respective physical limits. The exception is spectral efficiency which, at $\sim17\%$ below the physical limit, shows larger potential for further improvements, which is important, given that efforts to improve white LED performance across different efficiency loss channels through R\&D continue \cite{cho2017white, Weisbuch2020}.

@@

    \caption*{Note: Inv. - Year of initial invention, identified by the earliest literature source describing the original invention or idea in a field outside white LED lighting that eventually ‘spilled over’ to the latter field. S/O - Year of spillover to LED; Comm. - Year of commercial application, identified as the year of the first recorded application of that idea or invention in a commercial LED product. Ref. - References. LED innovations are ordered by the year in which a technology spillover into LED occurred, provided in the S/O column. Origin column represents knowledge domains in which spillovers initially emerged, where (S) denotes a scientific discipline and (T) is an area of technology. Ref. column lists literature sources for the represented discoveries, innovations and spillovers. Area of Improvement column represents the impact of spillovers on different aspects of white LED technology, e.g., improvements in particular sub-efficiencies.}

@@

While additional work is necessary to determine if the patterns we find for innovation in white LEDs apply in other demand side technologies, there are some emerging findings that would have implications. First, the fact that efficiency channels with greater physical understanding were more likely to be shaped by spillovers could indicate that investing into research that develops such physical understanding could reduce the cost of entry for new researchers and inventors and lead to spillovers and faster innovation. This could be more important for small-scale consumer facing technologies or products (e.g., cars, refrigerators, windows, food) that may face lower barriers to entry in research compared to, for instance, nuclear power or offshore wind. Second, we also find that innovation focused on consumer performance was incredibly important for this demand side consumer-facing technology and that the improvement mostly relied on integrating external knowledge, suggesting that a joint focus on cost and consumer experience in R\&D is both necessary but also requires a particularly high level of external knowledge integration, and possibly wider hiring practices. This finding also suggests that innovation in demand-side technologies, which have been according to some scholars ‘marginalized’ \cite{Wilson2012}, should be further promoted. A third possible hypothesis to test with future work on other demand side technologies is related to potentially larger contribution of demand-pull policies on cost metrics over time through their impact of scale up and learning by doing compared to R\&D. If corroborated in other technologies, this finding would add to evidence on the major role of regulations, bans and incentives shaping technology costs and trajectories.”

@@

\section*{Contributions}
All authors designed the study. M.P.W. and S.K. carried out the main experiments under the supervision of L.D.A. The data was analysed by M.P.W. All authors contributed to the discussion of the results and commented on the manuscript. The draft was written by M.P.W. and S.K.

>>>>

Improvements in technology cost and performance, which in this study we generally call “innovations”, can occur via different mechanisms, including (but not limited to) learning by doing \cite{WRIGHT_1936, Arrow_1962}, targeted research and development (R\&D) efforts, or economies of scale \cite{johansson2012global, iea2020perspectives}. The role of these mechanisms at different stages of the technology life cycle, from research and development to demonstration, market formation and diffusion \cite{grubler2012policies}, is an area of active research \cite{kavlak2018evaluating,Ziegler2021}. Different mechanisms can result in different kinds of innovations, characterized by what changes in the technology (e.g., its architecture, materials, components, or manufacturing process) and the scale of resulting improvements (e.g., radical or incremental) \cite{Acemoglu_2022}. For example, R\&D efforts (sometimes also called “learning-by-researching”) \cite{Castelnuovo_2005} more often than other mechanisms have been associated with substantial or even radical improvements in technology performance. On the other hand, learning-by-doing is typically associated with continuous incremental improvements in the technology resulting from accumulated experience in technology demonstration, manufacturing, and use. Both mechanisms, as well as economies of scale, can also lead to radical (i.e., discontinuous) or incremental reductions in technology costs.

@@

We find that total LED device efficiency of such LEDs increased from 5.8\% to 38.8\% between 2003 and 2020. A majority of this increase can be attributed to a series of innovations in white LED technology resulting from R\&D efforts, with at least 8.5\% of the cumulative increase attributed specifically to technological innovations driven by technology spillovers. Our interviews also revealed a lesser but still important role of incremental manufacturing process improvements most likely resulting from learning-by-doing. However, substantial further research is needed to quantify the relative contribution of learning-by-doing vis-a-vis R\&D to the efficiency increase during that time period. On the other hand, we find that innovations driven by technology spillovers contributed to nearly all consumer experience improvements in white LED lighting technology, thus playing an important role in widespread adoption of SSL in general illumination applications. Finally, our bottom-up LED manufacturing cost model shows a 95.5\% decrease in the cost of producing low-to-mid-power classic-chip-architecture GaN-based phosphor-converted warm-white LEDs from 1.11\$ to 0.05\$ (in 2020 USD) between 2003 and 2020. In contrast with LED performance improvements, where progress has been driven mostly by R\&D efforts, the dramatic decline in LED manufacturing cost has been a product of learning-by-doing resulting in higher yields across manufacturing steps and economies of scale resulting from increases in the sapphire wafer size. We provide the details of our findings for the progress in white LED efficiency, cost, and consumer experience, as well as the role of technology spillovers in this progress, in the remaining sections below.

@@

In \cref{fgr:breakthroughs_efficiency} we also compare, for the first time, efficiency improvements across sub-efficiencies over time, contrasting them with the physical limits of the corresponding loss channels. Since we focus on highest performing devices, and because we lack reliable information about the distribution of input parameters, which would require us to arbitrarily assign range values for inputs, these estimates do not show uncertainty ranges. Thus, our point estimates should be interpreted as the expected values in a best case scenario. We find that there has been consistent progress across all device sub-efficiencies in the recorded period. Specifically, between 2003 and 2020, forward voltage efficiency increased from 70\% to 99.5\%, internal quantum efficiency from 55\% to 90\%, electrical droop from 65\% to 90\%, light extraction efficiency from 60\% to 90\%, spectral efficiency from 74\% to 83\%, conversion efficiency (red) from 11\% to 45\%, conversion efficiency (green) from 19\% to 61\%. Notably, some sub-efficiencies for the most recent devices considered in our study are now within $\sim10\%$ of their respective physical limits. The exception is spectral efficiency which, at $\sim17\%$ below the physical limit, shows larger potential for further improvements, which is important, given that efforts to improve white LED performance across different efficiency loss channels through R\&D continue \cite{cho2017white, Weisbuch2020}.

@@

    \caption*{Note: Inv. - Year of initial invention, identified by the earliest literature source describing the original invention or idea in a field outside white LED lighting that eventually ‘spilled over’ to white LED lighting. S/O - Year of spillover to LED; Comm. - Year of commercial application, identified as the year of the first recorded application of that idea or invention in a commercial LED product. Ref. - References. LED innovations are ordered by the year in which a technology spillover into LED occurred, provided in the S/O column. Origin column represents knowledge domains in which spillovers initially emerged, where (S) denotes a scientific discipline and (T) is an area of technology. Ref. column lists literature sources for the represented discoveries, innovations and spillovers. Area of Improvement column represents the impact of spillovers on different aspects of white LED technology, e.g., improvements in particular sub-efficiencies.}

@@

While additional work is necessary to determine if the patterns we find for innovation in white LEDs apply in other demand side technologies, there are some emerging findings that would have implications. First, the fact that efficiency channels with greater physical understanding were more likely to be shaped by spillovers could indicate that investing into research that develops such physical understanding could reduce the cost of entry for new researchers and inventors and lead to spillovers and faster innovation. This could be more important for small-scale consumer facing technologies or products (e.g., cars, refrigerators, windows, food) that may face lower barriers to entry in research compared to, for instance, nuclear power or offshore wind. Second, we also find that innovation focused on consumer performance was incredibly important for this demand side consumer-facing technology and that the improvement mostly relied on integrating external knowledge, suggesting that a joint focus on cost and consumer experience in R\&D is both necessary but also requires a particularly high level of external knowledge integration, and possibly wider hiring practices. This finding also suggests that innovation in demand-side technologies, which have been according to some scholars ‘marginalized’ \cite{Wilson2012}, should be further promoted. A third possible hypothesis to test with future work on other demand side technologies is related to potentially larger contribution of demand-pull policies on cost metrics over time through their impact of scale up and learning by doing compared to R\&D. If corroborated in other technologies, this finding would add to evidence on the major role of regulations, bans and incentives shaping technology costs and trajectories.

Finally, the lessons learned about the mechanisms and drivers of rapid progress in white LED technology may have broader implications both for other demand-side or energy-efficiency technologies, and for low-carbon energy technologies with similar characteristics such as the granularity of technology, its modularity, and complexity \cite{malhotra2020accelerating, Wilson2012}. By comparing the role of these factors at a granular level across different low-carbon and energy efficiency technologies, we can identify and generalize key cross-technology patterns or differences that would help us formulate recommendations for industry and policymakers aimed at accelerating further clean energy innovation for climate change mitigation.

@@

\section*{Contributions}
All authors designed the study. M.P.W. and S.K. carried out the main empirical work under the supervision of L.D.A. Performance simulation software code was written by M.P.W. Data was analyzed by M.P.W. and S.K. All authors contributed to the writing of the manuscript.
